Introducing Statistical Design of Experiments to SPARQL Endpoint Evaluation

### 2013-05-08 exp1/iswc2013.tex
remove LenthPlot
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@533 43ef1190-90fd-4dd7-a769-d0121f02b80e
Fix all needed references
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@535 43ef1190-90fd-4dd7-a769-d0121f02b80e

These fragments has been carefully designed for illustrative utility,
as well as suitable selectivity. They are not themselves important,
they only serve to illustrate certain examples of possible
factors. When the experiment is run, they are combined to yield a
query, which is then sent to a SPARQL endpoint. The SPARQL endpoint is
itself set up using 4store\todo{ref} version 1.1.5. The ``Machine''
factor specifies a hostname, whereas ``TripleC'' and ``Implement''
specifies a port number. On each host, four 4store instances runs
independently on different ports. \todo{possible problem?}

@@

The experimentation is implemented in R\todo{ref}, which is a free
software environment for statistical computing and graphics. Necessary
tools for DoE has been implemented by the R community in packages
called DoE.base, BsMD and FrF2. The experiments are run on a third
computer in a 1 gigabit Ethernet LAN with the two experiment hosts.  As
response variable, we have chosen to use the time from the namespace
server lookup finishes to the data has been transferred. This choice is
somewhat arbitrary, many other response variables could be chosen, and
indeed, future work should explore multi-variable responses, but for
simplicity, we think this reflects the total amount of work done by
the SPARQL endpoint well. For the measurements, we have chosen to use
the RCurl package, which is a thin wrapper around the curl
library. Curl has well developed facilities for timing requests and
responses, and so we rely on its measurements.

@@

To proceed beyond a visual judgment of significance, we may use
Lenth's method\todo{ref Lenth}, which is a simple method for
estimating significance in an experiment with no replicates. We plot
the results in figure~\ref{fig:fulllenth}\todo{relationship level-SME/ME}. In
figure~\ref{fig:fullnormal}, Lenth's method is used to label the
significant effects. Visually, this is not as useful, since the large
number\todo{really include?} We see that we have very significant
effects. We also note that ``Implement'' itself is significantly
negative. We also note that the ``Implement:Lang'' interaction is
significantly positive, as expected from our setup, where ``Lang'' was
the detrimental side-effect we simulated.

\begin{figure}[h]
  \includegraphics[width=\textwidth]{fulllenth.pdf}
  \caption{A Lenth plot of the Full Factorial Experiment. The
    conservative Lenth criterion is indicated by the horizontal lines
    labelled SME, the liberal criterion is labelled ME. They are at
    ±0.90 and ±0.46 respectively. }\label{fig:fulllenth}
\end{figure}

@@

\bibliographystyle{plain}
%\bibliographystyle{abbrv}
%\bibliographystyle{jbact}
%\bibliographystyle{splncs03}
\bibliography{selectivity,federation,benchmarks,optimization,experimentsperformance}

>>>>

These fragments has been carefully designed for illustrative utility,
as well as suitable selectivity. They are not themselves important,
they only serve to illustrate certain examples of possible
factors. When the experiment is run, they are combined to yield a
query, which is then sent to a SPARQL endpoint. The SPARQL endpoint is
itself set up using 4store (see \cite{harris20094store}) version
1.1.5. The ``Machine'' factor specifies a hostname, whereas
``TripleC'' and ``Implement'' specifies a port number. On each host,
four 4store instances runs independently on different
ports. \todo{possible problem?}

@@

The experimentation is implemented in R\cite{rmanual}, which is a free
software environment for statistical computing and graphics. Necessary
tools for DoE has been implemented by the R community in packages
called DoE.base\cite{doebase} and FrF2\cite{frf2}. The experiments are run on a third
computer in a 1 gigabit Ethernet LAN with the two experiment hosts.  As
response variable, we have chosen to use the time from the namespace
server lookup finishes to the data has been transferred. This choice is
somewhat arbitrary, many other response variables could be chosen, and
indeed, future work should explore multi-variable responses, but for
simplicity, we think this reflects the total amount of work done by
the SPARQL endpoint well. For the measurements, we have chosen to use
the RCurl package\cite{lang2007r}, which is a thin wrapper around the curl
library. Curl has well developed facilities for timing requests and
responses, and so we rely on its measurements.

@@

To proceed beyond a visual judgment of significance, we may use
Lenth's method\cite{lenthmethod}, which is a simple method for
estimating significance in an experiment with no replicates. In
figure~\ref{fig:fullnormal}, Lenth's method is used to label the
significant effects. Visually, this is not as useful, since the large
number\todo{how to fix this...?} We see that we have very significant
effects. We also note that ``Implement'' itself is significantly
negative. We also note that the ``Implement:Lang'' interaction is
significantly positive, as expected from our setup, where ``Lang'' was
the detrimental side-effect we simulated.

@@

\bibliographystyle{plain}
%\bibliographystyle{abbrv}
%\bibliographystyle{jbact}
%\bibliographystyle{splncs03}
\bibliography{practical,selectivity,federation,benchmarks,optimization,experimentsperformance}
